Community-conditioned language models for linguistic variation

### 2021-10-13 paper/paper1.tex
anicdotal evidence figure & discussion

To this end, we compare the CCLM-learned community
embedding with the community embedding created by \citet{Kumar2018},
which were generated using using a negative-sampling optimization
algorithm, with the message author-community co-occurrence matrix as ground
truth, using data from January 2014 to April 2017.
We refer the reader
to \citet{Kumar2018}
for details, but the important point is that no linguistic
information is used to create these embeddings: they only reflect the social
relationship between communities, via 
the user-community membership relation. In contrast, CCLM community embeddings depend in no
way on which user is the author of any given message: we only use the
contents of messages, not authorship data.
%

@@

To get a handle on the relationship between the linguistic and social embeddigs

\subsection{Comparing embeddings: statistical evidence}

@@

Let us call \(L\) the matrix of (normed) linguistic embeddings and
\(S\) the matrix of (normed) social embeddings. Thus \(L_i\) and
\(S_i\) are the linguistic and social embeddings for community \(i\).
%
A straightforward, but ultimately flawed, generalization of the
methodology used to collect data in \cref{sec:storytelling} is to
consider each pair of communities \((i,j)\), and compute the
correlation between the cosine similarities of both embeddings. That
is, compute the Pearson correlation factor of the data set:

\[C = \{(x = L_i · L_j, y = S_i · S_j) \text {~for~} i,j ∈ [1,512]\}\] % BN: changed \C to C. Definition for \C may be missing

One could then use a standard statistical test to verify to obtain a
p-value for the correlation [TODO: what do we get?].  The flaw of
this methodology is that the number of pairs grows with the square
of the number of communities. With 510 communities, we have 129795
pairs, and standard statistical tests on Pearson correlation will
assure that we have a statistically significant correlation in all
but the weakest correlation. A further flaw is that the distribution of data points
of \(C\) are \emph{not} independent. Far from it, the are generated
from 1/510 independent variables. We consider this last flaw fatal, and take the following
approach instead.

@@

Instead of comparing embedding pairs, we will compare embeddings
community by community.

At this stage, one might naively think that it would suffice to
calculate the distance between two embeddings index-wise, which is
equal to the Frobenius distance between \(L\) and \(S\):

@@

The problem with this approach is that even if several dimensions of
\(L\) and \(S\) are correlated, they will not coincide in the
\emph{representation}. That is, applying a simple rotation (orthogonal
transformation) on either matrix widely changes the \(||L-S||_F\)
correlation metric. Another way to consider the same problem is that,
applying such a rotation does not change the cosine similarities
considered in the previous section.

@@

Here, the orthogonal matrix \(Ω\) gives a map from linguistic embeddings to social embeddings.
The problem of computing \(d(L,S)\) is known as the orthogonal Procrustes problem~\citep{Gower2004}. The
solution is
\[d(L,S) = n - \mathsf {Tr} (Σ)\] where the matrix \(Σ\) is obtained by the singular value
decomposition (SVD) \(U^TΣV = LS^T\). Additionally the vectors of \(U\) and \(V\) give
the directions of correlation respectively of \(L\) and \(S\). That is, each
singular value \(\sigma_i\) (the elements of the diagonal matrix \(Σ\)), give a measure of how much
correlation there is between the directions \(U_i\) and \(V_i\).

@@

\input{floats/pca_aligned}
\begin{figure*}[t]
    \centering
    \PCAAligned 
  \caption{First two components of the aligned social (left) and linguistic (right) embeddings,
    where the lingusitic embedding is taken from the LTSM with $l_c=1$.
    Correlation between these directions is given by $\sigma_0 = 53.4$ and $\sigma_1 = 35.6$.
    Colors are assigned by k-means clustering of the social embedding. 
    This figure is reproduced in the supplementary materials with a legend that helps 
    characterise the clusters.
  }
  \label{fig:pca-aligned}
\end{figure*}

>>>>

To this end, we compare the CCLM-learned community
embeddings\footnote{
  In this section, we only consider the embeddings from 
  the \emph{best} (highest informaiton gain) CCLM from each architecture family; 
  that is, the LSTM with $l_c=1$ and the Transformer with $l_c=0$, however we
  observed similar results for other values of $l_c$.}
with the community embedding created by \citet{Kumar2018},
which were generated using using a negative-sampling optimization
algorithm, with the message author-community co-occurrence matrix as ground
truth, using data from January 2014 to April 2017.
We refer the reader
to \citet{Kumar2018}
for details, but the important point is that no linguistic
information is used to create these embeddings: they only reflect the social
relationship between communities, via 
the user-community membership relation. In contrast, CCLM community embeddings depend in no
way on which user is the author of any given message: we only use the
contents of messages, not authorship data.
%

@@

\begin{figure*}[ht]
\begin{tikzpicture}
  \begin{groupplot}[group style={
      group size=2 by 1,
      horizontal sep=1cm,
    },
    height=5cm,width=8cm,
    axis x line*=bottom, axis y line*=left,x axis line style={opacity=0},
    ymin=-0.35, ymax=1.00,
    xtick=data,
    xticklabel style={align=left, anchor=east, rotate=60},
    minor tick length=1ex,
    ]
    \nextgroupplot[ybar,
      bar width=0.25cm,
      xticklabels from table={floats/low_soc_high_ling.csv}{comms-label},
      ylabel={cosine similarity},
    ]
    \addplot[draw=violet,fill=violet!10] table [y=lstm-sim] {floats/low_soc_high_ling.csv};
    \addplot[draw=teal,fill=teal!10] table [y=transformer-sim] {floats/low_soc_high_ling.csv};
    \addplot[draw=red,fill=red!10] table [y=social-sim] {floats/low_soc_high_ling.csv};
    \nextgroupplot[ybar,
      bar width=0.25cm,
      xticklabels from table={floats/low_ling_high_soc.csv}{comms-label},
      legend style={font=\tiny,at={(1,1)},anchor=south east}, legend columns=3
    ]
    \addplot[draw=violet,fill=violet!10] table [y=lstm-sim] {floats/low_ling_high_soc.csv};
    \addplot[draw=teal,fill=teal!10] table [y=transformer-sim] {floats/low_ling_high_soc.csv};
    \addplot[draw=red,fill=red!10] table [y=social-sim] {floats/low_ling_high_soc.csv};
    \legend{LSTM, Transformer, Social}
  \end{groupplot}
\end{tikzpicture}
\caption{
  Communities with high linguistic similarity but low social similarity (left) and \textit{vice versa} (right). 
  Linguistic similarity is measured with the embeddings from the best LSTM ($l_c=1$) and Transformer ($l_c=0$)
  models and social similarity is measured with the embedding from \citet{Kumar2018}.
}
\label{fig:embedding-differences-examples}
\end{figure*}

Let us call \(L\) the matrix of (normed) linguistic embeddings and
\(S\) the matrix of (normed) social embeddings. Thus \(L_i\) and
\(S_i\) are the linguistic and social embeddings for community \(i\).

A straightforward, but ultimately flawed, way to measure 
how similar the two spaces are is to
consider each pair of communities \((i,j)\), and compute the
correlation between the cosine similarities of both embeddings. That
is, compute the Pearson correlation factor of the data set:
\[C = \{(x = L_i · L_j, y = S_i · S_j) \text {~for~} i,j ∈ [1,510]\}\] % BN: changed \C to C. Definition for \C may be missing

Indeed, $C$ shows positive correlation for both the LSTM ($r=0.438$)
and transformer ($r=0.452$) linguistic embeddings.\footnote{
  By comparison, the correlation between the two linguistic embeddings is \num{0.759}.}
The correlations are significant with $p<0.001$ in all cases;
however, we note that the number of pairs grows with the square
of the number of communities (with 510 communities, we have 129795)
pairs), meaning that standard statistical tests on Pearson correlation will
assure us of statistical significance in all
but the weakest of correlations. A further flaw is that the data points in 
\(C\) are \emph{not} distributed independently---far from it since each data point is generated 
from \num{2} of \num{510} independent variables. We consider this last flaw fatal, 
and take a different approach for computing the correlation between community
embeddings (\cref{sec:statistical-evidence}).

We can, however, use cosine similarity to find pairs of communities
that are similar in one embedding but different in the other. 
This gives us a way of characterizing the differences between the two kinds of
community representation.
The examples shown in \cref{fig:embedding-differences-examples} were selected by taking the \num{5} 
most linguistically similar (in the LSTM embedding) pairs of communities that had below median social similarity 
(and \textit{vice versa}).\footnote{
Median similarity among pairs of communities was \num{0.177} for the social embedding and
\num{0.010} and \num{0.012} for the LSTM and transformer linguistic embeddings, respectively.}

As one would expect, it seems that the communities with similar linguistic embeddings
but dissimilar social embeddings (\cref{fig:embedding-differences-examples}, left) share a similar
topic, mode of interaction, or language variety, but for various reasons, may attract a quite
different set of members.
For example, \subreddit{hiphopheads} and \subreddit{Monstercat} are both topically related to music,
but the music genres are different, and the later has a more geographically local focus (Monstercat is an 
independent electronic music label based in Vancouver).
While both \subreddit{MLPLounge} and \subreddit{CasualConversation} are oriented towards casual general-topic conversation,
the former is geared towards members of a niche internet sub-culture. % adult fans of the television show My Little Pony. There are other subreddits (mentioned in r/MLPLounge's "about") that are for on-topic discussion of MLP
Both \subreddit{rocketbeans} and \subreddit{de} are primarily German-language subreddits, but the former is comprised of 
fans of a computer gaming-related YouTube channel, while the later is more general-interest.

Differences at the other end of the spectrum (\cref{fig:embedding-differences-examples}, right)
are somewhat harder to interpret.
It is mostly easy to see why these communities would have different linguistic embeddings---%
in all cases the topics are quite different.
The reason they have similar social embeddings is less obvious,
but we can discern some trends in how the communities are premised.
The \subreddit{progresspics} and \subreddit{TalesFromRetail} are premised, in part,
on seeking support from other people with similar experiences;
\subreddit{legaladvice}, \subreddit{Cooking}, and \subreddit{loseit}
all have a knowledge-sharing component to them;
\subreddit{running} and \subreddit{Coffee} are hobby-focused;
and \subreddit{self} (often) and \subreddit{askscince} (by premise) are
places people ask and answer questions.
It may be that there are different patterns in the \emph{social function}
that people attribute to this particular social media website---%
people who use Reddit in one way are more likely
to belong to communities that are premised on the same kind of social function,
even if the topics (and indeed language) of those communities are quite different.
Testing this hypothesis would require a more focused study design and 
ideally consider communities from multiple social networks (online or otherwise).

\subsection{Comparing embeddings: statistical evidence}
\label{sec:statistical-evidence}

@@

@@

Instead of comparing embedding pairs, as in \cref{sec:storytelling}, 
we will compare embeddings community by community.
A naive approach would be to 
calculate the distance between two embeddings index-wise, which is
equal to the Frobenius distance between \(L\) and \(S\):

@@

The problem with this approach is that even if several dimensions of
\(L\) and \(S\) are correlated, they will not coincide in the
\emph{representation}. That is, applying a simple rotation (orthogonal
transformation) on either matrix widely changes the \(||L-S||_F\)
correlation metric. Another way to consider the same problem is that,
applying such a rotation does \emph{not} change the cosine similarities
considered in the previous section.

@@

Here, the orthogonal matrix \(Ω\) gives a map from linguistic embeddings to social embeddings.
The problem of computing \(d(L,S)\) is known as the orthogonal Procrustes problem~\citep{Gower2004}.\footnote{
This approach has also been used to compare word embeddings across representations \citep[e.g.,][]{Hamilton2016}.}
The solution is
\[d(L,S) = n - \mathsf {Tr} (Σ)\] where the matrix \(Σ\) is obtained by the singular value
decomposition (SVD) \(U^TΣV = LS^T\). Additionally the vectors of \(U\) and \(V\) give
the directions of correlation respectively of \(L\) and \(S\). That is, each
singular value \(\sigma_i\) (the elements of the diagonal matrix \(Σ\)), give a measure of how much
correlation there is between the directions \(U_i\) and \(V_i\).

@@

\input{floats/pca_aligned}
\begin{figure*}[t]
    \centering
    \PCAAligned 
  \caption{First two components of the aligned social (left) and linguistic (right) embeddings,
    where the lingusitic embedding is taken from the LTSM with $l_c=1$.
    Correlation between these directions is given by $\sigma_0 = 53.4$ and $\sigma_1 = 35.6$.
    Colors are assigned by k-means clustering of the social embedding. 
    This figure is reproduced in the supplementary materials with a legend that helps 
    to characterise the clusters.
  }
  \label{fig:pca-aligned}
\end{figure*}
